(Deep) Induction for GADTs

### 2021-04-12 induction-paper/ENTCS-GADT-induction.tex
Revise Section 4.4

We now present an induction rule for the $\mathsf{LTerm}$ type defined in Example~\ref{example:lterm}.
For that, we need liftings for both $\mathsf{LType}$ and
$\mathsf{LTerm}$. We will also need a lifting of predicates for arrow types, since arrow types appear in 
$\mathsf{LType}$ and $\mathsf{LTerm}$. The lifting for arrow types 
\[
  \mathsf{Arr^{\wedge} : \forall (A\, B : Set) \to (A \to Set) \to (B \to Set) \to (A \to B) \to Set}
\]
is defined as
\[
  \mathsf{Arr^{\wedge}\, A\, B\, Q_A\, Q_B\, f = \forall (a : A) \to Q_A\,a \to Q_B\, (f\,a)}
\]
Now we can define the lifting for $\mathsf{LType}$, which has type 
\[
\mathsf{LType^{\wedge} : \forall (A : Set) \to (A \to Set) \to LType\,A \to Set}
\]
and is defined by pattern-matching as
\[
  \mathsf{LType^{\wedge}\,A\,Q_A\,(tBool\,B\,e)
  = \sum [Q_B : B \to Set]\, Equal^{\wedge}\, A\, B\, Q_A\, K^{Bool}_{\top} \,e}
\]
where $\mathsf{Q_A : A \to Set}$,
$\mathsf{K^{Bool}_{\top}}$ is the constantly $\mathsf{\top}$-valued predicate on $\mathsf{Bool}$,
$\mathsf{e : Equal\,A\,Bool}$,
and
\[
\begin{array}{l}
\mathsf{LType^{\wedge}\,A\,Q_A\,(tArr\, B\, C\, e\, T_B\, T_C)} \\
\quad
\mathsf{= \sum [Q_B : B \to Set] [Q_C : C \to Set]\,
  Equal^{\wedge}\,A\, (B \to C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C) \, e
  \times \, LType^{\wedge}\,B\,Q_B\,T_B \times LType^{\wedge}\,C\,Q_C\,T_C
}
\end{array}
\]
where $\mathsf{Q_A : A \to Set}$, $\mathsf{e : Equal\, A\, (B \to C)}$, $\mathsf{T_B : LType \,B}$, and 
$\mathsf{T_C : LType \,C}$,
and 
\[
\mathsf{LType^{\wedge}\,A\,Q_A\,(tList\, B\, e\, T_B)
=
\sum [Q_B : B \to Set]\, 
  Equal^{\wedge}\,A\, (List\, B)\, Q_A\, (List^{\wedge} \, B\, Q_B) \, e 
  \times LType^{\wedge}\,B\,Q_B\,T_B
}
\]
where $\mathsf{Q_A : A \to Set}$, $\mathsf{e : Equal\, A\, (List\, B)}$, and $\mathsf{T_B : LType \,B}$.

Given the lifting for $\mathsf{LType}$, we can define the lifting for $\mathsf{LTerm}$. Again, 
the lifting 
\[
\mathsf{LTerm^{\wedge} : \forall (A : Set) \to (A \to Set) \to LTerm\,A \to Set}
\]
is defined by pattern-matching as
\[
  \mathsf{LTerm^{\wedge}\,A\,Q_A\,(var\,s\,T_A) = LType^{\wedge}\, A\, Q_A\, T_A}
\]
where $\mathsf{s : String}$ and $\mathsf{T_A : LType\, A}$, and
\[
\begin{array}{l}
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (abs \,B \,C \,e \,s \,T_B \,t_C)} \\ 
\quad \mathsf{
  = \sum [Q_B : B \to Set] [Q_C : C \to Set]\,
      Equal^{\wedge} \, A\, (B \to C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C)\, e
      \times \, LType^{\wedge}\, B\, Q_B\, T_B
      \times \, LTerm^{\wedge}\, C\, Q_C\, t_C }
\end{array}
\]
where $\mathsf{e : Equal \, A \, (B \to C)}$,
$\mathsf{s : String}$,
$\mathsf{T_B : LType\,B}$,
and $\mathsf{t_C : LTerm \, C}$,
and
\[
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (app\, B\, t_{BA}\, t_B)
=
\sum [Q_B : B \to Set]\, 
  LTerm^{\wedge}\, (B \to A)\, (Arr^{\wedge} \, B\, A\, Q_B \, Q_A)\, t_{BA}
  \times LTerm^{\wedge}\, B\, Q_B\, t_B}
\]
where $\mathsf{t_{BA} : LTerm \, (B \to A)}$ and $\mathsf{t_B : LTerm \, B}$, and
\[
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (listC\, B\, e\, ts)
=
\sum [Q_B : B \to Set]\,
 Equal^{\wedge} \, A\, (List\,B)\, Q_A\, (List^{\wedge} \, B\, Q_B) \, e
 \times List^{\wedge}\, (LTerm\,B) \, (LTerm^{\wedge} \, B\, Q_B) \, ts}
\]
where $\mathsf{e : Equal\, A\, (List \,B)}$ and $\mathsf{ts : List\, (LTerm B)}$. 
Notice we use the lifting for $\mathsf{List}$ in the case for the $\mathsf{listC}$ constructor.

With these liftings, we can define the deep induction principle for $\mathsf{LTerm}$.
% Var case 
Let $\mathsf{dIndVar}$ be the function 
\begin{multline*}
\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A \to Set)} \\
\mathsf{\to 
  \forall (A : Set) (Q_A : A \to Set) (s : String) (T_A : LType\, A) 
  \to LType^{\wedge} \, A\, Q_A\, T_A
  \to P \, A\, Q_A\, (var \, s\, T_A)
  }
\end{multline*}
associated to the $\mathsf{var}$ constructor. 
% Abs case 
Let $\mathsf{dIndAbs}$ be the function 
\begin{align*}
  &\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A \to Set)} \\
  &\quad\mathsf{\to 
  \forall (A\,B\,C: Set) (Q_A : A \to Set)  (Q_B : B \to Set) (Q_C : C \to Set)
  (e : Equal\, A\, (B \to C)) (s : String) } \\ 
  &\quad\mathsf{ \to (T_B : LType\, B) \to (t_C : LTerm\, C)
  \to Equal^{\wedge}\,A\,(B \to C)\, Q_A \, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C) \, e
  } \\
  &\quad\mathsf{
  \to LType^{\wedge}\, B\, Q_B\, T_B
  \to P\, C\, Q_C\, t_C\, 
  \to P \, A\, Q_A\, (abs \,B \,C \, e \,s \,T_B \, t_C)
  }
\end{align*}
associated to the $\mathsf{abs}$ constructor. 
% App case 
Let $\mathsf{dIndApp}$ be the function 
\begin{align*}
  &\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A \to Set)} \\
  &\quad \mathsf{\to 
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
   (t_{BA} : LTerm\, (B \to A)) (t_B : LTerm\, B)} \\
  &\quad \mathsf{
  \to P\, (B \to A)\, (Arr^{\wedge} \, B\, A\, Q_B \, Q_A) \, t_{BA} \, 
  \to P\, B\, Q_B\, t_B\, 
  \to P \, A\, Q_A\, (app \,B \,t_{BA} \, t_B) }
\end{align*}
associated to the $\mathsf{app}$ constructor. 
% ListC case 
Let $\mathsf{dIndListC}$ be the type 
\begin{align*}
  &\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A \to Set)} \\
  &\quad \mathsf{\to 
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
    (e : Equal\, A\, (List\, B)) (ts : List\, (LTerm\, B))} \\ 
  &\quad \mathsf{
    \to Equal^{\wedge}\, A\, (List\,B)\, Q_A\, (List^{\wedge}\, B\, Q_B)\, e 
  \to List^{\wedge}\, (LTerm\,B) (P\, B\, Q_B)\, ts
  \to P \, A\, Q_A\, (listC \,B \,e \, ts) }
\end{align*}
associated to the $\mathsf{listC}$ constructor.

@@

For the $\mathsf{abs}$ case, let $\mathsf{t_A = abs \,B \,C \, e \,s \,T_B \, t_C)} $ and define

@@

where $\mathsf{(Q_B , lift_{t_{BA}}, list_{t_B}) : LTerm^{\wedge}\, A\, Q_A\, (app \,B \,t_{BA} \, t_B)}$

@@

  p_{ts} = dIndLTerm\, P\, cvar\, cabs\, capp\, clistc\, B\, Q_B :  PredMap \,(LTerm^{\wedge}\, B\, Q_B) \, (P\,B\,Q_B)

@@

  \to PredMap\,Q_A\,Q'_A \to PredMap \,(List^{\wedge}\, A\, Q_A)\, (List^{\wedge}\, A\, Q'_A)}

@@

$\mathsf{PredMap}$ gives the type of morphisms of predicates and is defined as: 

@@

  &\mathsf{PredMap : \forall\, \{A : Set\} \to (A \to Set) \to (A \to Set) \to Set } \\
  &\mathsf{PredMap \, Q_A\,Q'_A\, = \forall\, (a : A) \to Q_A\,a \to Q'_A\,a}

@@

\mathsf{data\, G\,(\ol{a : Set}) : Set\,where}\\
\mathsf{\;\;\;\;\;\;\;\;C_i\, :\, F_i\,G\,\ol{b} \to G (\ol{K_i\,\ol{b}})}

@@

is defined by a finite number of constructors $\mathsf{C_i}$.

@@

$\mathsf{\ol{a}}$ is a list of types of length $\mathsf{\alpha}$,
so that it can be applied to the type constructor $\mathsf{G}$ of arity $\mathsf{\alpha}$.

@@

Each of the $\mathsf{\alpha}$-many $\mathsf{K_i}$ is a type constructor of arity $\mathsf{\beta}$
so that it can be applied to the list of types $\mathsf{\ol{b}}$ of length $\mathsf{\beta}$.

@@

Moreover, notice that the arity of $\mathsf{G}$ matches the number of type constructors $\mathsf{\ol{K_i}}$.

@@

\item $\mathsf{F_i\,G\,\ol{b} = G(\ol{F_a\,\ol{b}})}$
where none of the $\mathsf{F_a}$ contains the recursive variable,
i.e., $\mathsf{F_a : Set^\beta \to Set}$

@@

for each $\mathsf{a}$.

@@

\item $\mathsf{F_i\,G\,\ol{b} = H\,\ol{b}}$

@@

so, in particular, the unit and empty types, $\mathsf{1}$ and $\mathsf{0}$.

@@

\item $\mathsf{F_i\,G\,\ol{b} = H (\ol{F_c\,G\,\ol{b}})}$

@@

and $\mathsf{F_c}$ has the same signature as $\mathsf{F_i}$  and is built recursively from the same induction rules, for every $\mathsf{c = 1 \dots \gamma}$.

@@

This way $\mathsf{H}$ admits functorial semantics~\cite{jp20}, and thus we have a map function for $\mathsf{H^{\wedge}}$,

@@

\mathsf{HLMap : \forall (\ol{c : Set}) (\ol{Q_c\;Q_c' : c \to Set})
	\to \ol{PredMap\,c\,Q_c\,Q_c'}
	\to PredMap\,(H\,\ol{c})\,(H^{\wedge}\,\ol{c}\,\ol{Q_c})\,(H^{\wedge}\,\ol{c}\,\ol{Q_c'})}

@@

where $\mathsf{PredMap : \forall (c : Set) \to (c \to Set) \to (c \to Set) \to Set}$
is defined as
\mathsf{PredMap\,c\,Q_c\,Q_c' = \forall (x : c) \to Q_c\,x \to Q_c'\,x}
and represents the type of morphisms between predicates.

@@

\mathsf{F_i\,G\,\ol{b} :=
F_i'\,G\,\ol{b} \times F_i''\,G\,\ol{b}
\ \vert\ F_i'\,G\,\ol{b} + F_i''\,G\,\ol{b}
\ \vert\ F_i'\,\ol{b} \to F_i''\,G\,\ol{b}
\ \vert\ G(\ol{F_a\,\ol{b}})
\ \vert\ H\,\ol{b}
\ \vert\ H (\ol{F_c\,G\,\ol{b}})

@@

Consider a generic GADT as defined above,
\begin{equation}\label{eq:gadt}
\begin{array}{l}
\mathsf{data\, G\,(a : Set) : Set\,where}\\
\mathsf{\;\;\;\;\;\;\;\;C\, :\, F\,G\,\ol{b} \to G (K\,\ol{b})}
\end{array}
\end{equation}
which, for ease of notation,
we assume to be a unary type constructor (i.e. it depends on a single type parameter $\mathsf{a}$)
and to have only one constructor $\mathsf{C}$.
Extending the argument to GADTs of arbitrary arity and with multiple constructors presents no difficulty
other than heavier notation.
In the definition above, $\mathsf{F}$ has signature $\mathsf{(Set \to Set) \to Set^{\beta} \to Set}$
and each $\mathsf{K}$ has signature $\mathsf{Set^{\beta} \to Set}$.
The constructor $\mathsf{C}$ can be rewritten using the $\mathsf{Equal}$ type as
\[
\mathsf{C : \exists (\ol{b : Set}) \to Equal\,a\,(K\,\ol{b}) \to F\,G\,\ol{b} \to G\,a}
\]
which is the form we shall use from now on.

In order to state the induction rule for $\mathsf{G}$,
we first need to define $\mathsf{G}$'s associated predicate-lifting function 
\[
\mathsf{G^{\wedge} : \forall (a : Set) \to (a \to Set) \to G\,a \to Set}
\]
as
\[
\mathsf{G^{\wedge}\,a\,Q_a\,(C\,\ol{b}\,e\,x)
= \exists (\ol{Q_b : b \to Set})
\to Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e
\times F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,{\ol{Q_b}}\,x}
\]
where $\mathsf{Q_a : a \to Set}$
and $\mathsf{C\,\ol{b}\,e\,x : G\,a}$,
i.e., $\mathsf{e : Equal\,a\,(K\,\ol{b})}$ and $\mathsf{x : F\,G\,\ol{b}}$.
As already mentioned before, we also assume to have liftings for $\mathsf{F}$,
\begin{multline*}
\mathsf{F^{\wedge} : \forall (G : Set^{\alpha} \to Set) (\ol{b : Set})
\to (\forall (a : Set) \to (a \to Set) \to G\,a \to Set)} \\
\mathsf{\to (\ol{b \to Set})
\to F\,G\,\ol{b} \to Set}
\end{multline*}
and for  $\mathsf{K}$,
\[
\mathsf{K^{\wedge} : \forall (\ol{b : Set}) \to (\ol{b \to Set}) \to K\,\ol{b} \to Set}
\]
Finally, associate the function
\begin{multline*}
\mathsf{CC = \lambda (P : \forall (a : Set) \to (a \to Set) \to G\,a \to Set)} \\
\mathsf{\to \forall (a : Set) (\ol{b : Set}) (Q_a : a \to Set) (\ol{Q_b : b \to Set}) (e : Equal\,a\,(K\,\ol{b})) (x : F\,G\,\ol{b})} \\
\mathsf{\to Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e
	\to F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}\,x
	\to P\,a\,Q_a\,(C\,\ol{b}\,e\,x)}
\end{multline*}
to the constructor $\mathsf{C}$.

@@

\mathsf{\forall (P : \forall (a : Set) \to (a \to Set) \to G\,a \to Set) \to CC\,P} \\

@@

\mathsf{\to \forall (a : Set)(Q_a : a \to Set)(y : G\,a) \to G^{\wedge}\,a\,Q_a\,y \to P\,a\,Q_a\,y}

@@

we validate the induction rule by providing a term $\mathsf{DIG}$ for the type above.

@@

\mathsf{DIG\,P\,cc\,a\,Q_a\,(C\,\ol{b}\,e\,x)\,(\ol{Q_b}, L_E, L_F)
= cc\,a\,\ol{b}\,Q_a\,\ol{Q_b}\,e\,x\,L_E\,(p\,x\,L_F)}

@@

where $\mathsf{cc : CC\,P}$ and

@@

\item $\mathsf{C\,\ol{b}\,e\,x : G\,a}$,
i.e., $\mathsf{e : Equal\,a\,(K\,\ol{b})}$,
and $\mathsf{x : F\,G\,\ol{b}}$;
\item $\mathsf{(\ol{Q_b}, L_E, L_F) : G^{\wedge}\,a\,Q_a (C\,\ol{b}\,e\,x)}$,

@@

$\mathsf{Q_b : b \to Set}$ for each $\mathsf{b}$,
$\mathsf{L_E : Equal^{\wedge}\,a\,(K\,\ol{b})\,Q_a\,(K^{\wedge}\,\ol{b}\,\ol{Q_b})\,e}$,
and $\mathsf{L_F : F^{\wedge}\,G\,\ol{b}\,G^{\wedge}\,{\ol{Q_b}}\,x}$.

@@

\mathsf{F^{\wedge}\,G\,\ol{b}\,P\,\ol{Q_b}

>>>>

We now present an induction rule for the $\mathsf{LTerm}$ type defined in Example~\ref{example:lterm}.
For that, we need liftings for both $\mathsf{LType}$ and
$\mathsf{LTerm}$. We will also need a lifting of predicates for arrow types, since arrow types appear in 
$\mathsf{LType}$ and $\mathsf{LTerm}$. The lifting for arrow types 
\[
  \mathsf{Arr^{\wedge} : \forall (A\, B : Set) \to (A \to Set) \to (B \to Set) \to (A \to B) \to Set}
\]
is defined as
\[
  \mathsf{Arr^{\wedge}\, A\, B\, Q_A\, Q_B\, f = \forall (a : A) \to Q_A\,a \to Q_B\, (f\,a)}
\]
Now we can define the lifting for $\mathsf{LType}$, which has type 
\[
\mathsf{LType^{\wedge} : \forall (A : Set) \to (A \to Set) \to LType\,A \to Set}
\]
and is defined by pattern-matching as
\[
  \mathsf{LType^{\wedge}\,A\,Q_A\,(tBool\,B\,e)
  = \sum [Q_B : B \to Set]\, Equal^{\wedge}\, A\, B\, Q_A\, K^{Bool}_{\top} \,e}
\]
where $\mathsf{Q_A : A \to Set}$,
$\mathsf{K^{Bool}_{\top}}$ is the constantly $\mathsf{\top}$-valued predicate on $\mathsf{Bool}$
and $\mathsf{e : Equal\,A\,Bool}$;
\[
\begin{array}{l}
\mathsf{LType^{\wedge}\,A\,Q_A\,(tArr\, B\, C\, e\, T_B\, T_C)} \\
\quad
\mathsf{= \sum [Q_B : B \to Set] [Q_C : C \to Set]\,
  Equal^{\wedge}\,A\, (B \to C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C) \, e
  \times \, LType^{\wedge}\,B\,Q_B\,T_B \times LType^{\wedge}\,C\,Q_C\,T_C
}
\end{array}
\]
where $\mathsf{Q_A : A \to Set}$, $\mathsf{e : Equal\, A\, (B \to C)}$, $\mathsf{T_B : LType \,B}$ and 
$\mathsf{T_C : LType \,C}$;
and 
\[
\mathsf{LType^{\wedge}\,A\,Q_A\,(tList\, B\, e\, T_B)
=
\sum [Q_B : B \to Set]\, 
  Equal^{\wedge}\,A\, (List\, B)\, Q_A\, (List^{\wedge} \, B\, Q_B) \, e 
  \times LType^{\wedge}\,B\,Q_B\,T_B
}
\]
where $\mathsf{Q_A : A \to Set}$, $\mathsf{e : Equal\, A\, (List\, B)}$, and $\mathsf{T_B : LType \,B}$.
%Given the lifting for $\mathsf{LType}$, 
We can define the lifting for $\mathsf{LTerm}$. Again, 
the lifting 
\[
\mathsf{LTerm^{\wedge} : \forall (A : Set) \to (A \to Set) \to LTerm\,A \to Set}
\]
is defined by pattern-matching as
\[
  \mathsf{LTerm^{\wedge}\,A\,Q_A\,(var\,s\,T_A) = LType^{\wedge}\, A\, Q_A\, T_A}
\]
where $\mathsf{s : String}$ and $\mathsf{T_A : LType\, A}$;
\[
\begin{array}{l}
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (abs \,B \,C \,e \,s \,T_B \,t_C)} \\ 
\quad \mathsf{
  = \sum [Q_B : B \to Set] [Q_C : C \to Set]\,
      Equal^{\wedge} \, A\, (B \to C)\, Q_A\, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C)\, e
      \times \, LType^{\wedge}\, B\, Q_B\, T_B
      \times \, LTerm^{\wedge}\, C\, Q_C\, t_C }
\end{array}
\]
where $\mathsf{e : Equal \, A \, (B \to C)}$,
$\mathsf{s : String}$,
$\mathsf{T_B : LType\,B}$,
and $\mathsf{t_C : LTerm \, C}$;
\[
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (app\, B\, t_{BA}\, t_B)
=
\sum [Q_B : B \to Set]\, 
  LTerm^{\wedge}\, (B \to A)\, (Arr^{\wedge} \, B\, A\, Q_B \, Q_A)\, t_{BA}
  \times LTerm^{\wedge}\, B\, Q_B\, t_B}
\]
where $\mathsf{t_{BA} : LTerm \, (B \to A)}$ and $\mathsf{t_B : LTerm \, B}$; and
\[
\mathsf{LTerm^{\wedge}\,A\,Q_A\, (listC\, B\, e\, ts)
=
\sum [Q_B : B \to Set]\,
 Equal^{\wedge} \, A\, (List\,B)\, Q_A\, (List^{\wedge} \, B\, Q_B) \, e
 \times List^{\wedge}\, (LTerm\,B) \, (LTerm^{\wedge} \, B\, Q_B) \, ts}
\]
where $\mathsf{e : Equal\, A\, (List \,B)}$ and $\mathsf{ts : List\, (LTerm B)}$. 
Notice that we use the lifting for $\mathsf{List}$ in the case for the $\mathsf{listC}$ constructor.

With these liftings, we can define the deep induction principle for $\mathsf{LTerm}$.
% Var case 
Let $\mathsf{dIndVar}$ be the function 
\begin{multline*}
\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A \to Set)} \\
\mathsf{\to 
  \forall (A : Set) (Q_A : A \to Set) (s : String) (T_A : LType\, A) 
  \to LType^{\wedge} \, A\, Q_A\, T_A
  \to P \, A\, Q_A\, (var \, s\, T_A)
  }
\end{multline*}
associated to the $\mathsf{var}$ constructor. 
% Abs case 
Let $\mathsf{dIndAbs}$ be the function 
\begin{align*}
  &\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A \to Set)} \\
  &\quad\mathsf{\to 
  \forall (A\,B\,C: Set) (Q_A : A \to Set)  (Q_B : B \to Set) (Q_C : C \to Set)
  (e : Equal\, A\, (B \to C)) (s : String) } \\ 
  &\quad\mathsf{ \to (T_B : LType\, B) \to (t_C : LTerm\, C)
  \to Equal^{\wedge}\,A\,(B \to C)\, Q_A \, (Arr^{\wedge} \, B\, C\, Q_B \, Q_C) \, e
  } \\
  &\quad\mathsf{
  \to LType^{\wedge}\, B\, Q_B\, T_B
  \to P\, C\, Q_C\, t_C\, 
  \to P \, A\, Q_A\, (abs \,B \,C \, e \,s \,T_B \, t_C)
  }
\end{align*}
associated to the $\mathsf{abs}$ constructor. 
% App case 
Let $\mathsf{dIndApp}$ be the function 
\begin{align*}
  &\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A \to Set)} \\
  &\quad \mathsf{\to 
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
   (t_{BA} : LTerm\, (B \to A)) (t_B : LTerm\, B)} \\
  &\quad \mathsf{
  \to P\, (B \to A)\, (Arr^{\wedge} \, B\, A\, Q_B \, Q_A) \, t_{BA} \, 
  \to P\, B\, Q_B\, t_B\, 
  \to P \, A\, Q_A\, (app \,B \,t_{BA} \, t_B) }
\end{align*}
associated to the $\mathsf{app}$ constructor. 
% ListC case 
Let $\mathsf{dIndListC}$ be the function 
\begin{align*}
  &\mathsf{\lambda (P : \forall (A : Set) \to (A \to Set) \to LTerm\,A \to Set)} \\
  &\quad \mathsf{\to 
  \forall (A \,B : Set) (Q_A : A \to Set)  (Q_B : B \to Set) 
    (e : Equal\, A\, (List\, B)) (ts : List\, (LTerm\, B))} \\ 
  &\quad \mathsf{
    \to Equal^{\wedge}\, A\, (List\,B)\, Q_A\, (List^{\wedge}\, B\, Q_B)\, e 
  \to List^{\wedge}\, (LTerm\,B) (P\, B\, Q_B)\, ts
  \to P \, A\, Q_A\, (listC \,B \,e \, ts) }
\end{align*}
associated to the $\mathsf{listC}$ constructor.

@@

For the $\mathsf{abs}$ case, let $\mathsf{t_A = abs \,B \,C \, e \,s \,T_B \, t_C} $ and define

@@

where $\mathsf{(Q_B , lift_{t_{BA}}, list_{t_B}) : LTerm^{\wedge}\, A\, Q_A\, (app \,B \,t_{BA} \, t_B)}$, i.e.,

@@

  p_{ts} = dIndLTerm\, P\, cvar\, cabs\, capp\, clistc\, B\, Q_B :  PredMap\,(LTerm\,B) \,(LTerm^{\wedge}\, B\, Q_B) \, (P\,B\,Q_B)

@@

  \to PredMap\,A\,Q_A\,Q'_A \to PredMap\,(List\,A) \,(List^{\wedge}\, A\, Q_A)\, (List^{\wedge}\, A\, Q'_A)}

@@

$\mathsf{PredMap}$ is the type constructor that produces the type of morphisms between predicates, and it is defined as follows: 

@@

  &\mathsf{PredMap : \forall\, (A : Set) \to (A \to Set) \to (A \to Set) \to Set } \\
  &\mathsf{PredMap \,A\, Q_A\,Q'_A\, = \forall\, (a : A) \to Q_A\,a \to Q'_A\,a}

@@

\mathsf{data\ G : Set^\alpha \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;c_i\, :\, F_i\,G\,\ol{B} \to G (\ol{K_i\,\ol{B}})}

@@

is defined by a finite number of constructors $\mathsf{c_i}$.

@@

%$\mathsf{\ol{a}}$ is a list of types of length $\mathsf{\alpha}$,
%so that it can be applied to the type constructor $\mathsf{G}$ of arity $\mathsf{\alpha}$.

@@

each of the $\mathsf{\alpha}$-many $\mathsf{K_i}$ is a type constructor of arity $\mathsf{\beta}$
so that it can be applied to the list of types $\mathsf{\ol{B}}$ of length $\mathsf{\beta}$.

@@

Moreover, notice that the arity of $\mathsf{G}$ matches the number of type constructors $\mathsf{K_i}$.

@@

\item $\mathsf{F_i\,G\,\ol{B} = G(\ol{F_j\,\ol{B}})}$
where none of the $\mathsf{\alpha}$-many $\mathsf{F_j}$ contains the recursive variable,
i.e., $\mathsf{F_j : Set^\beta \to Set}$

@@

for each $\mathsf{j = 1, \dots, \alpha}$.

@@

\item $\mathsf{F_i\,G\,\ol{B} = H\,\ol{B}}$

@@

so, in particular, the unit and empty types, $\mathsf{1}$ and $\mathsf{0}$,
and the case in which $\mathsf{F_i}$ is a variable.

@@

\item $\mathsf{F_i\,G\,\ol{B} = H (\ol{F_k\,G\,\ol{B}})}$

@@

and $\mathsf{F_k}$ has the same signature as $\mathsf{F_i}$  and is built recursively from the same induction rules, for every $\mathsf{k = 1 \dots \gamma}$.

@@

This way we know that $\mathsf{H}$ admits functorial semantics~\cite{jp20}, and thus there is a map function for $\mathsf{H^{\wedge}}$,

@@

\mathsf{HLMap : \forall (\ol{C : Set}) (\ol{Q_C\;Q_C' : C \to Set})
	\to \ol{PredMap\,C\,Q_C\,Q_C'}
	\to PredMap\,(H\,\ol{C})\,(H^{\wedge}\,\ol{C}\,\ol{Q_C})\,(H^{\wedge}\,\ol{C}\,\ol{Q_C'})}

@@

%where $\mathsf{PredMap : \forall (C : Set) \to (C \to Set) \to (C \to Set) \to Set}$
%is defined as
%\[
%\mathsf{PredMap\,C\,Q_c\,Q_c' = \forall (x : c) \to Q_c\,x \to Q_c'\,x}
%\]
%and represents the type of morphisms between predicates.

@@

\mathsf{F_i\,G\,\ol{B} :=
F_i'\,G\,\ol{B} \times F_i''\,G\,\ol{B}
\ \vert\ F_i'\,G\,\ol{B} + F_i''\,G\,\ol{B}
\ \vert\ F_i'\,\ol{B} \to F_i''\,G\,\ol{B}
\ \vert\ G(\ol{F_j\,\ol{B}})
\ \vert\ H\,\ol{B}
\ \vert\ H (\ol{F_k\,G\,\ol{B}})

@@

Consider a generic GADT as defined above,
\begin{equation}\label{eq:gadt}
\begin{array}{l}
\mathsf{data\ G : Set \to Set\ where}\\
\mathsf{\;\;\;\;\;\;\;\;c\, :\, F\,G\,\ol{B} \to G (K\,\ol{B})}
\end{array}
\end{equation}
which, for ease of notation,
we assume to be a unary type constructor (i.e. it depends on a single type parameter $\mathsf{A}$)
and to have only one constructor $\mathsf{c}$.
Extending the argument to GADTs of arbitrary arity and with multiple constructors presents no difficulty
other than heavier notation.
In the definition above, $\mathsf{F}$ has signature $\mathsf{(Set \to Set) \to Set^{\beta} \to Set}$
and each $\mathsf{K}$ has signature $\mathsf{Set^{\beta} \to Set}$.
The constructor $\mathsf{c}$ can be rewritten using the $\mathsf{Equal}$ type as
\[
\mathsf{c : \forall (\ol{B : Set}) \to Equal\,A\,(K\,\ol{B}) \to F\,G\,\ol{B} \to G\,A}
\]
which is the form we shall use from now on.

In order to state the induction rule for $\mathsf{G}$,
we first need to define $\mathsf{G}$'s associated predicate-lifting function 
\[
\mathsf{G^{\wedge} : \forall (A : Set) \to (A \to Set) \to G\,A \to Set}
\]
as
\[
\mathsf{G^{\wedge}\,A\,Q_A\,(c\,\ol{B}\,e\,x)
= \sum [\ol{Q_B : B \to Set}]\,
Equal^{\wedge}\,A\,(K\,\ol{B})\,Q_A\,(K^{\wedge}\,\ol{B}\,\ol{Q_B})\,e
\times F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,{\ol{Q_B}}\,x}
\]
where $\mathsf{Q_A : A \to Set}$
and $\mathsf{c\,\ol{B}\,e\,x : G\,A}$,
i.e., $\mathsf{e : Equal\,A\,(K\,\ol{B})}$ and $\mathsf{x : F\,G\,\ol{B}}$.
As already mentioned before, we also assume to have liftings for $\mathsf{F}$,
\[
\mathsf{F^{\wedge} : \forall (G : Set^{\alpha} \to Set) (\ol{B : Set})
\to (\forall (A : Set) \to (A \to Set) \to G\,A \to Set)
\to (\ol{B \to Set})
\to F\,G\,\ol{B} \to Set}
\]
and for  $\mathsf{K}$,
\[
\mathsf{K^{\wedge} : \forall (\ol{B : Set}) \to (\ol{B \to Set}) \to K\,\ol{B} \to Set}
\]
Finally, associate the function
\[
\begin{array}{l}
\mathsf{dIndC = \lambda (P : \forall (A : Set) \to (A \to Set) \to G\,A \to Set)} \\
\quad \mathsf{\to \forall (A : Set) (\ol{B : Set}) (Q_A : A \to Set) (\ol{Q_B : B \to Set}) (e : Equal\,A\,(K\,\ol{B})) (x : F\,G\,\ol{B})} \\
\quad \mathsf{\to Equal^{\wedge}\,A\,(K\,\ol{B})\,Q_A\,(K^{\wedge}\,\ol{B}\,\ol{Q_B})\,e
	\to F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}\,x
	\to P\,A\,Q_A\,(c\,\ol{B}\,e\,x)}
\end{array}
\]
to the constructor $\mathsf{c}$.

@@

\mathsf{\forall (P : \forall (A : Set) \to (A \to Set) \to G\,A \to Set) \to dIndC\,P

@@

\to \forall (A : Set)(Q_A : A \to Set)(y : G\,A) \to G^{\wedge}\,A\,Q_A\,y \to P\,A\,Q_A\,y}

@@

we validate the induction rule by providing a term $\mathsf{dIndG}$ for the type above.

@@

\mathsf{dIndG\,P\,cc\,A\,Q_A\,(c\,\ol{B}\,e\,x)\,(\ol{Q_B}, liftE, liftF)
= cc\,A\,\ol{B}\,Q_A\,\ol{Q_B}\,e\,x\,liftE\,(p\,x\,liftF)}

@@

where $\mathsf{cc : dIndC\,P}$ and

@@

\item $\mathsf{c\,\ol{B}\,e\,x : G\,A}$,
i.e., $\mathsf{e : Equal\,A\,(K\,\ol{B})}$,
and $\mathsf{x : F\,G\,\ol{B}}$;
\item $\mathsf{(\ol{Q_B}, liftE, liftF) : G^{\wedge}\,A\,Q_A (c\,\ol{B}\,e\,x)}$,

@@

$\mathsf{Q_B : B \to Set}$ for each $\mathsf{B}$,
$\mathsf{liftE : Equal^{\wedge}\,A\,(K\,\ol{B})\,Q_A\,(K^{\wedge}\,\ol{B}\,\ol{Q_B})\,e}$,
and $\mathsf{liftF : F^{\wedge}\,G\,\ol{B}\,G^{\wedge}\,{\ol{Q_B}}\,x}$.

@@

\mathsf{F^{\wedge}\,G\,\ol{B}\,P\,\ol{Q_B}
